Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-02-09 draft.tex

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/SN_efficiencies.png}
    \caption{SN detection efficiencies split up by DES field. The $y-$axis represents the fraction of simulated SNe in a given redshift bin that would have been detected by DES-SN and passed light curve quality cuts. The curves are capped at efficiencies around 0.5 because DES is restricted to $\sim6$ month observing seasons each calendar year.
    \label{fig:SN_efficiency}}
\end{figure}

@@

To correct for these incompleteness, we follow a similar method to that used in the PTF rates analyses of \citet{Frohmaier2019,Frohmaier2020} as well as the DES-SN superluminous SN rate of \citet{Thomas2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate 230,000 SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated with SEDs based upon the SALT2 model \citep{Guy2007} and trained on the Joint Lightcurve Analysis dataset \citep{Betoule2014}. The SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on location, explosion epoch, and redshift, which we define as the SN efficiency. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i}) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i}\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i}\right)}\right)\,.
\end{equation}

The distribution of efficiency as a function of redshift is shown in Figure \ref{fig:SN_efficiency}. It is evident that the deep fields (X3, C3) are sensitive to SNe at higher redshifts, while there is no drastic shifts between efficiencies in the eight shallow fields.

@@

\subsection{Second order processes affecting the supernova rate \label{subsec:discussion}}

@@

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/SN_efficiencies.png}
    \includegraphics[width=0.5\textwidth]{figs/SN_efficiencies.png}
    \caption{SN detection efficiencies split up by DES field. The $y-$axis represents the fraction of simulated SNe in a given redshift bin that would have been detected by DES-SN and passed light curve quality cuts. The curves are capped at efficiencies around 0.5 because DES is restricted to $\sim6$ month observing seasons each calendar year.
    \label{fig:SN_efficiency}}
\end{figure}

@@

To correct for these incompleteness, we follow a similar method to that used in the PTF rates analyses of \citet{Frohmaier2019,Frohmaier2020} as well as the DES-SN superluminous SN rate of \citet{Thomas2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate 230,000 SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated with SEDs based upon the SALT2 model \citep{Guy2007} and trained on the Joint Lightcurve Analysis dataset \citep{Betoule2014}. The SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on a combination of sky location, explosion epoch, redshift, stretch, and colour. The fraction of recovered SNe (the efficiency) is thus described by a 5-dimensional surface. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, with stretch $x_1$ and colour $c$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_1,c) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i},x_1,c\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i},x_{1,i},c_i\right)}\right)\,.
\end{equation}

In practice, the gradient of the efficiency function is strongest between different DES fields and as a function of redshift, while SN stretch and colour have smaller effects. We integrate the efficiencies across the full simulated time range, and as such the efficiency is limited to $\sim 0.5$ due to the 6-month nature of the DES observing seasons.
The distribution of efficiency as a function of redshift is shown in Figure \ref{fig:SN_efficiency}. It is evident that the deep fields (X3, C3) are sensitive to SNe at higher redshifts, while there is no drastic shifts between efficiencies in the eight shallow fields.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_x1.png}
    \caption{As per Fig. \ref{fig:rate_fitted} but with the data split by their stretch parameter $x_1$.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_x1}}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm_Qerf1.1_split_c.png}
    \caption{As per Fig. \ref{fig:rate_fitted} but with the data split by their colour parameter $c$.%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fitted_split_c}}
\end{figure}

\renewcommand{\arraystretch}{1.2}
\begin{table}
	\centering
	\caption{Results of the Bayesian parameter estimation for the SN Ia DTD.}
	\label{tab:dtd_results_split_lc}
	\begin{tabular}{crr} % four columns, alignment for each
		\hline
		 Sample &$\beta$ & $A$ \\
		 &-       & $10^{-13}$ M$_{\odot}^{-1}$yr$^{-1}$ \\
		\hline
		Fiducial & $-1.18\pm0.09$ &  $1.16\pm0.08$ \\
	    $x_1 < 0$ & $-0.79\pm0.08$ &  $1.19 \pm0.05$ \\
		$x_1 > 0$ & $-1.68 _{ -0.15} ^{ +0.14}$ & $0.51\pm0.12$ \\
		$c < 0$ & $-1.09 _{ -0.07} ^{ +0.08}$ & $0.89\pm0.04$ \\
		$c > 0$ & $-1.19\pm0.09$ & $1.16\pm0.08$ \\
		\hline
	\end{tabular}
\end{table}
\subsection{Second order processes affecting the supernova rate \label{subsec:discussion}}

@@

\section{The delay-time distribution as a function of supernova properties}
\label{sec:split_x1_c}

In the previous sections we have considered all SNe that passed the various quality, standardisation, and redshift cuts. However, it is well established that SNe Ia show correlations of varying strength between properties of their light curves and host galaxies. For example, measures of the light curve duration (e.g. decline rate $\Delta m_{15}$, or stretch such as SALT2 $x_1$) are known to correlate with host galaxy properties such as stellar mass \citep{Lampeitl2010,Sullivan2010}.
